### 2024-04-18 other/datasheet/datasheet.qmd
finished datasheet

---
title: "Datasheet for 'Deaths of People Experiencing Homelessness' and 'Toronto Shelter System Flow'"
author: 
  - Maroosh Gillani
thanks: "Code and data are available at: https://github.com/Maroosh-Gillani/homeless_mortality"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The 'Deaths of People Experiencing Homelessness' dataset aims to track the deaths of Toronto's homeless population.
    - The 'Toronto Shelter System Flow' dataset is used to track the details regarding Toronto's homeless shelters and their users.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The 'Deaths of People Experiencing Homelessness' dataset was created by Toronto Public Health (TPH) in collaboration with The Shelter, Support and Housing Administration (SSHA) and health and social service agencies.
    - The 'Toronto Shelter System Flow' dataset was published by Toronto Shelter & Support Services, and the Shelter Management Information System (SMIS) is used to track the people using shelter services.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - 'Deaths of People Experiencing Homelessness': not funded.
    - 'Toronto Shelter System Flow': Funded by the City of Toronto.
4. *Any other comments?*
    - No.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- 'Deaths of People Experiencing Homelessness': Each row shows the number of homeless people who passed away each month, starting from January 2017 to June 2023. 
	- 'Toronto Shelter System Flow': 7 rows are dedicated to a single month (starting from January 2018). Of these seven rows, each row is dedicated to specific types of the population that is using Toronto's homeless shelters, such as Families, Youth, Refugees etc. For each different type of population, a count for how many fall into specific categories (such as newly identified, returned from housing, return to shelter, age groups, gender etc.) is provided.
2. *How many instances are there in total (of each type, if appropriate)?*
	- 'Deaths of People Experiencing Homelessness': There are 81 rows, and each row has 4 columns. All columns are numeric except for Month of death, which is made of characters.
	- 'Toronto Shelter System Flow': There are 556 rows, and each row has 18 columns. All columns are numeric except for date(mmm-yy) and population_group, which are characters.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
  - 'Deaths of People Experiencing Homelessness': The data is complete and contains all instances.
	- 'Toronto Shelter System Flow': The data is complete and contains all instances.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- 'Deaths of People Experiencing Homelessness': Each row shows the number of homeless people who passed away each month, starting from January 2017 to June 2023. 
	- 'Toronto Shelter System Flow': 7 rows are dedicated to a single month (starting from January 2018). Of these seven rows, each row is dedicated to specific types of the population that is using Toronto's homeless shelters, such as Families, Youth, Refugees etc. For each different type of population, a count for how many fall into specific categories (such as newly identified, returned from housing, return to shelter, age groups, gender etc.) is provided.
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- Yes, both datasets have ids for each row, stored in a designated id column.
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
		- 'Deaths of People Experiencing Homelessness': The data only contains deaths of homeless people that are reported to the TPH by the SSHA partners. The data is also subject to change based on additional reports to TPH.
	- 'Toronto Shelter System Flow': The dataset contains data from people who have used overnight services only, and is subject to change as new data becomes available.
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- 'Deaths of People Experiencing Homelessness': Yes, each row is dedicated to a specific month and year, and how many homeless deaths occurred in that month.
	- 'Toronto Shelter System Flow': Yes, each row is dedicated to a specific month, year and population group.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- No.
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- No.
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- Both datasets are self-contained.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- 'Toronto Shelter System Flow': Yes, there are 6 population groups (Chronic, Refugees, Families, Youth, Single Adult, Non-Refugees) which are further divided into 14 categories (returned_from_housing, returned_to_shelter, age16-24, gender_male etc.)
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- No.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- No.
16. *Any other comments?*
	- No.

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- Directly observable and collected from subjects (for homeless shelters through SMIS)
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
  - 'Deaths of People Experiencing Homelessness': TPH collects, analyzes and reports data. SSHA share information about death with TPH.
	- 'Toronto Shelter System Flow': Usage of SMIS software.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- Not a sample from larger dataset.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- 'Deaths of People Experiencing Homelessness': deceased homeless people of Toronto. Not compensated.
	- 'Toronto Shelter System Flow': Homeless Shelter Users of Toronto. Not compensated.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- 'Deaths of People Experiencing Homelessness': January 2017 to June 2023.
	- 'Toronto Shelter System Flow': January 2018 to February 2024.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No.
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- Obtained from Open Data Toronto.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- Data collection notification is subject to the policies of organizations collecting data.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- Data collection consent is subject to the policies of organizations collecting data.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- Data collection consent revocation is subject to the policies of organizations collecting data.
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- No.
12. *Any other comments?*
	- No.

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- Not to my knowledge.
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- https://open.toronto.ca/dataset/deaths-of-people-experiencing-homelessness/ and https://open.toronto.ca/dataset/toronto-shelter-system-flow/
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- No.
4. *Any other comments?*
	- No.

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- No.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- No.
3. *What (other) tasks could the dataset be used for?*
	- Both datasets can be used to analyze homeless populations, and factors affecting their mortality, and shelter usage trends.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- No.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- No.
6. *Any other comments?*
	- No.

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- No.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- Both datasets are available for public use through OpenDataToronto's portal.
3. *When will the dataset be distributed?*
	- They are available to use currently.
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- Both datasets are subject to Ontario Open Government License.
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- No.
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No.
7. *Any other comments?*
	- No.

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- 'Deaths of People Experiencing Homelessness': maintained by OpenDataToronto, TPH and SSHA.
	- 'Toronto Shelter System Flow': maintained by OpenDataToronto, and SMIS.
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
  - 'Deaths of People Experiencing Homelessness': EDAU@toronto.ca
	- 'Toronto Shelter System Flow': tsssdata@toronto.ca
3. *Is there an erratum? If so, please provide a link or other access point.*
	- No.
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- 'Deaths of People Experiencing Homelessness': updated by maintainers semi-annually on the OpenDataToronto portal.
	- 'Toronto Shelter System Flow': updated by maintainers monthly on the OpenDataToronto portal.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- No.
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- Both datasets have new data just appended upon previous instances.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- Both datasets are available to be downloaded in various formats for extension/augmentation purposes.
8. *Any other comments?*
	- No.

\newpage

# References